---
title: 'Coronavirus Time Series Analysis'
output:
  html_notebook: default
  word_document: default
  html_document:
    df_print: paged
---

# Introduction


The current date of this analysis is March 16, 2020.

All code and data can be found on [Github](https://github.com/howardnewyork/corona).

# Summary of Results


# Methodology

Software: I used [Greta](https://greta-stats.org/index.html) and [Greta GP](https://github.com/greta-dev/greta.gp) to conduct a Bayesian analysis.  

Models:  The *basic model* is binomial model with a uniform prior assumption on the risk probability.  For the youngest and very oldest ages there is very little data and this prior influences the results.

I developed a model combining all age groups and a model that analyzed each age group separately.  For the latter, I used two types of models:
1. The first model assumes each age category is independent from the others.
2. The second model assumes a correlation structure between each age category.  This was achieved by applying a latent Gaussian Process over the age-dependent risk parameters, using a RBF (radial basis function) kernel.  I tested two GP models.

# Recommended Actions

*Disclaimer*: I am not an epidemiologist so my analysis is very much one of an armchair scientist. However, I am an actuary, so I know a thing or two about risk rates.

The analysis suggests the following:


# Data Sources

[Johns Hopkins COID-19 Data Repository](https://github.com/CSSEGISandData/COVID-19/compare?expand=1)

# Initialize Data

```{r setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=TRUE, message=FALSE}
# Time Series Analysis of Survival from Breast Cancer


library(rstan)
options(mc.cores = min(8,parallel::detectCores()))
rstan_options(auto_write = TRUE)

library(ggplot2)
library(dplyr)
library(purrr)
library(tidyr)
library(readr)
library(stringr)
library (bayesplot)
library(lubridate)
library(corrplot)

# MCMC Pars
n_samples_base = 1000
warmup_base = 250
chains_base = 20

# Set Directories
working_dir = paste0(getwd(), "/")
data_dir = paste0(working_dir, "data/")
output_dir = paste0(working_dir, "output/")
stan_dir = paste0(working_dir, "stan/")

selected_countries = c("Germany", "France", "Spain", "China", "Australia", "US", "South Korea", "Italy",  "United Kingdom")
selected_countries = c("Italy", "US", "Australia", "South Korea", "China")
selected_countries = c("Italy", "US",  "South Korea", "China")
cutoff= 200

# Load Data
file_date = "2020_03_15"
corona_raw = read_csv(file = paste0(data_dir, paste0("time_series_19_covid_", file_date, ".csv")))
head(corona_raw, 10)

corona = corona_raw %>%
  mutate(country = if_else(country == "Korea, South", "South Korea", country))  %>%  
  select(-lat, -long, -province_state) %>%
  group_by(country) %>%
  summarise_all(sum) %>%
  gather(date, confirmed, -country) %>%
  group_by(country) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(growth = confirmed - lag(confirmed),
         growth_rate = log(confirmed / lag(confirmed))) %>%
  arrange(date) %>%
  mutate(growth_rate = ifelse(is.finite(growth_rate), growth_rate, NA))  


corona_align = corona %>%
  filter(country %in% selected_countries,
         confirmed >=cutoff) %>%
  ungroup() %>%
  group_by(country) %>%
  mutate(day = date-first(date)+1) %>%
  select(country,confirmed, growth,growth_rate, date, day) %>%
  arrange(day)
corona_align

corona 
corona %>%  filter(country=="China")
corona %>%  filter(country=="Australia", confirmed >=10) %>% select(date, confirmed, growth_rate, growth_rate) 

```



# Data Exploration




```{r}
ggplot(corona %>% filter(country %in% selected_countries)) +
  geom_line(aes(date,confirmed, color=country)) + 
  labs(title = "Confirmed Cases by Country", x="date", y="Confirmed Cases")+
   scale_y_continuous( labels = scales::comma)

ggplot(corona %>% filter(country %in% selected_countries)) +
  geom_line(aes(date,growth, color=country)) + 
  labs(title = "Growth in Confirmed Cases by Country", x="date", y="Growth in Confirmed Cases")+
   scale_y_continuous( labels = scales::comma)

corona %>%  filter(country %in% selected_countries, confirmed >=10) %>% select(date, confirmed, growth_rate, growth_rate) %>% 
ggplot() +
  geom_line(aes(date,growth_rate, color=country)) + 
  labs(title = "Growth in Confirmed Cases by Country", x="date", y="Daily Growth Rate")+
  scale_y_continuous( labels = scales::comma) +
  theme_bw() +
  coord_cartesian(ylim=c(-1, 3))


```

Correlation Analysis

```{r}



ggplot(corona_align ) +
  geom_line(aes(day,confirmed, color=country)) + 
  labs(title = "Aligned Confirmed Cases by Country", x=paste0(c("Day >= ", cutoff, " Cases"), collapse = ""), y="Confirmed Cases")+
   scale_y_continuous( labels = scales::comma)

ggplot(corona_align ) +
  geom_line(aes(day,growth, color=country)) + 
  geom_point(aes(day,growth, color=country, shape=country)) + 
  labs(title = "Aligned Growth in Confirmed Cases by Country", x=paste0(c("Day >= ", cutoff, " Cases"), collapse=""), y="Growth in Confirmed Cases")+
   scale_y_continuous( labels = scales::comma)

ggplot(corona_align) +
  geom_line(aes(day,growth_rate, color=country)) + 
  geom_point(aes(day,growth_rate, color=country, shape=country)) + 
  labs(title = "Aligned Growth Rate by Country", x=paste0(c("Day >= ", cutoff, " Cases"), collapse=""), y="Daily Growth Rate")+
  theme_bw() 



corona_cor = corona_align %>%
  select(day, country, growth_rate) %>%
  pivot_wider(names_from = country, values_from = growth_rate) %>%
  select(-day) %>% 
  cor(use = "pairwise.complete.obs")
corrplot(corona_cor[selected_countries, selected_countries])

corona_align %>%
  group_by(country) %>%
  group_map(~lm(data = .x,  formula =  growth_rate ~ day ))


ave_growth = corona_align %>%
  group_by(country) %>%
  summarize(mean_growth = mean(growth_rate, na.rm = T),
            median_growth = median(growth_rate, na.rm = T) )

ggplot(ave_growth %>% pivot_longer(names_to = "type", values_to = "value", -country)) + 
  geom_point(aes(country, value, color=type), size=2)+ 
  labs(title = "Mean and Median Growth") + 
  theme_bw()

M= lm(data = corona_align, formula =  growth_rate ~ day + country) 
summary(M)
M = lm(data = corona_align, formula =  growth_rate ~ day + confirmed + country) 
summary(M)

M = lm(data = corona_align, formula =  growth_rate ~  confirmed + country) 
summary(M)


corona_align = corona_align %>%
  filter(is.finite(growth_rate)) %>%
  mutate(smoothed = predict(loess(formula =  growth_rate ~ as.numeric(day), na.action = na.omit)))

corona_align

ggplot(corona_align) +
  geom_line(aes(day,smoothed, color=country)) + 
  geom_point(aes(day,smoothed, color=country, shape=country)) + 
  labs(title = "Smoothed Aligned Growth Rate by Country", x=paste0(c("Day >= ", cutoff, " Cases"), collapse=""), 
       y="Daily Growth Rate")+
  theme_bw() 


corona_cor2 = corona_align %>%
  select(day, country, smoothed) %>%
  pivot_wider(names_from = country, values_from = smoothed) %>%
  select(-day) %>% 
  cor(use = "pairwise.complete.obs")
corrplot(corona_cor2[selected_countries, selected_countries])

#corr2 = corona_cor2

```


# Stan GP Analysis


Prepare input

```{r}
country_count = length(unique(corona_align$country))
N = nrow(corona_align)
N_pred = as.numeric(country_count * (max(corona_align$day) - min(corona_align$day)+1))


stan_list = list(
  N1 = N,  # historic data
  #N2= N_pred,  # predicted points
  N2= N,  # predicted points
  D=1, # Dimension of input values
  dimCat = country_count, # dimension of number oif unique categorical values
  
  # Historic Data
  x1 = matrix(as.numeric(corona_align$day), ncol = 1) ,  # input values (N1 x D matrix from R)
  x_cat1= as.numeric(as.factor(corona_align$country)),  # input values containing a categorical value
  y1 = corona_align$growth_rate,  # output values 
  # Inputs for prediction
  
  # x2= matrix(rep(min(corona_align$day):max(corona_align$day), each = country_count), ncol=1),  # input values (N xD matrix from R)
  # x_cat2 =  rep(1:country_count, times = max(corona_align$day)),  # input values containing a categorical value
  x2 = matrix(as.numeric(corona_align$day), ncol = 1) ,  # input values (N1 x D matrix from R)
  x_cat2= as.numeric(as.factor(corona_align$country)),  # input values containing a categorical value

 # inputs for priors on parameters
 prior_eta = 2,
 prior_rho = c(5,5),
 prior_sigma = 1,
 prior_alpha =2
)
str(stan_list)
```



Run Stan Model

```{r}
mod = stan_model(file = paste0(stan_dir, "gp_mod1.stan"))
draws = sampling(object = mod, data = stan_list , chains = 8, iter = 1000, warmup=300)
#summary(draws, pars = c("eta", "rho", "sigma", "fStar"))
```


```{r}

country_names = levels(as.factor(corona_align$country))

preds = tibble(
  day = c(stan_list$x2),
  country = country_names[stan_list$x_cat2],
  growth_rate = summary(draws, pars = c("fStar"))$summary[,"mean"]
)


str(preds)

ggplot(preds)  +
  geom_line(aes(day, growth_rate, color=country)) + 
  geom_point(aes(day, growth_rate, color=country, shape=country))  +
  labs(title = paste(c("Daily Growth Rates of Infection following\n", cutoff, " confirmed infections"), collapse =""), x=paste0(c("Day >= ", cutoff, " Cases"), collapse="")) + 
  coord_cartesian(ylim=c(0, .5))
  

```



Correlation Analysis

```{r}
#summary(draws, pars = "lkj2")$summary

country_cor = matrix(summary(draws, pars = "lkj2")$summary[, "mean"], nrow = country_count)
rownames(country_cor) = country_names
colnames(country_cor) = country_names
country_cor

```





# Stan Model with Block Diagonal

```{r}
mod_block = stan_model(file = paste0(stan_dir, "gp_mod2.stan"))

draws_block = sampling(object = mod_block, data = stan_list , chains = 8, iter = 1000, warmup=300)
summary(draws, pars = c("eta", "rho", "sigma", "fStar"))
```


```{r}

country_names = levels(as.factor(corona_align$country))

preds_block = tibble(
  day = c(stan_list$x2),
  country = country_names[stan_list$x_cat2],
  growth_rate = summary(draws_block, pars = c("fStar"))$summary[,"mean"]
)


str(preds_block)

ggplot(preds_block)  +
  geom_line(aes(day, growth_rate, color=country)) + 
  geom_point(aes(day, growth_rate, color=country, shape=country))  +
  labs(title = paste(c("Daily Growth Rates of Infection following\n", cutoff, " confirmed infections"), collapse =""), x=paste0(c("Day >= ", cutoff, " Cases"), collapse="")) + 
  coord_cartesian(ylim=c(0, .5))
  

```





### Greta Projection Analysis


#### Beta Binom Model: Single Parameter,

```{r}

# Identify the data input for Greta using the as_data function

library(onehot)

encoder = (onehot(as.data.frame(corona_align$country)))
predict(encoder,as.data.frame(corona_align$country))

country = as.matrix(predict(encoder,as.data.frame(corona_align$country)))
country_num = 1:country_count
country_num = country %*% country_num

index = matrix(1:(country_count^2), country_count, country_count)
country_rho_index = matrix(0, N, N)
for (i in 1:N){
  for (j in 1:N){
    country_rho_index[i,j] = index[country_num[i],country_num[j]]
  }
}
country_rho_index = c(country_rho_index)

growth_rate = as_data(corona_align$growth_rate)
day = as_data(as.numeric(corona_align$day))
country = as_data(country) # one hot encoded
country_num = as_data(country_num)

# Define the prior for the theta parameter
eta = lognormal(0,4)
rho_day = lognormal(0,4)
rho_country = lkj_correlation(eta = 1, dimension = country_count) *2 -1 

rho_country_long = zeros(country_count^2, 1)
rho_country_long = rho_country

rho_country_index = rho_country_long[country_rho_index]

rho_country_kernel = zeros(N,N)
rho_country_kernel = rho_country_index

# kernel & GP
kernel = rbf(rho_day, eta_deaths) 
f = gp(ages, kernel)  
theta_deaths = ilogit(f) # convert to unit interval

# Define the likehood for the model
#distribution(infected) = binomial(size = exposure, prob = theta_infected, dim = categories)  
#distribution(symptomatic) = binomial(size = exposure, prob = theta_symptomatic, dim = categories)  
#distribution(asymptomatic) = binomial(size = exposure, prob = theta_asymptomatic, dim = categories)
distribution(deaths) = binomial(size = infected, prob = theta_deaths, dim = categories)  

```

We can now establish and compile the model:

```{r results="hide"}
# Establish the model
#m_gp= model(theta_deaths, eta_deaths, rho_deaths)
m_gp= model(theta_deaths, eta_deaths) 

n_samples = n_samples_base; chains=chains_base; warmup=warmup_base*4

S = n_samples * chains # Total number of simulations
draws_gp=mcmc(m_gp, n_samples = n_samples, warmup = warmup,chains = chains)

```


```{r}
# diagnostics
theta_names = colnames(draws_gp[[1]])
mcmc_trace(draws_gp, pars = theta_names[1:10])
mcmc_trace(draws_gp, pars = theta_names[11:11])
mcmc_intervals(draws_gp, pars = theta_names[1:10])
mcmc_intervals(draws_gp, theta_names[11:11])

summary(draws_gp)

```


```{r}
theta_post_gp = 
  draws_gp %>% 
  reduce(rbind) %>%
  as_tibble() %>%
  select(-eta_deaths) 
  
colnames(theta_post_gp) = 
  paste0(rep(c( "theta_deaths_"),
             each=categories) , 1:categories)
theta_post_gp = cbind(S=1:S,as_tibble(theta_post_gp)) %>%
  gather(posterior, value, -S) %>%
  mutate(risk_cat = as.factor(gsub(pattern = "_([^_]*)$","",(posterior))), 
           age_cat = as.factor((as.numeric(extract_numeric(posterior)))*10-5))

theta_post_gp_sum = theta_post_gp %>%
  group_by(age_cat) %>%
  summarize(mean = mean(value),
            low = quantile(value, 0.1),
            high = quantile(value, 0.9)
            ) %>%
  gather(range, value, -mean, -age_cat)

ggplot(theta_post_gp_sum) +
  geom_point(aes(age_cat, mean, color = age_cat),size=2) +
  geom_line(aes(age_cat, value, group = age_cat)) +
  labs(title="Diamond Princess: Risk of Death", color="", y="Fatality Rate", x="Age") + 
  theme_bw()



# ggplot(theta_post_gp ) +
#   geom_violin(aes(age_cat, value, fill = age_cat), alpha=0.5, draw_quantiles = c(0.1, 0.5, 0.9), scale = "width") +
#   labs(title="Diamond Princess: Risk of Death", color="", y="Fatality Rate", x="Age") + 
#   ylim(0,.1)+
#   theme_bw()


ggplot(theta_post_gp_sum) +
  geom_point(aes(age_cat, mean, color = age_cat),size=2) +
  geom_line(aes(age_cat, value, group = age_cat)) +
  labs(title="Diamond Princess: Risk of Death\n with 80% Credible Interval", color="", y="Fatality Rate", x="Age") + 
  theme_bw()

gp_par =
  draws_gp %>% 
  reduce(rbind) %>%
  as_tibble() %>%
  select(eta_deaths) %>%
  gather(par, value)

ggplot(gp_par ) +
  geom_density(aes(value, fill=par), alpha=0.5) +
  labs(title="GP Pars") + 
  facet_wrap(~par, scales = "free")+
  theme_bw()


```

Summary Statistics

```{r}
risk_band_dependent = theta_post_gp %>%
  group_by(age_cat) %>%
  summarize(mortality_mean = mean(value),
            mortality_low = quantile(value, probs = .1),
            mortality_high = quantile(value, probs = .9))

gp_mortality_fixed_rho = ggplot(risk_band_dependent %>% mutate(age= as.numeric(as.character(age_cat)))) +
  geom_line(aes(age, mortality_mean)) + 
  geom_ribbon(aes(age, ymin=mortality_low, ymax=mortality_high), fill= "blue", alpha=.2)+
  theme_bw() + 
  labs(title="GP Model - Fatality Rate: 80% Credible Interval")+
  scale_x_continuous(breaks = diamond$age)
gp_mortality_fixed_rho

risk_band_dependent

```





#### Gaussian Process Model - Variable Rho

```{r}

# Identify the data input for Greta using the as_data function

categories = nrow(diamond)
exposure = as_data(diamond$exposure)
ages = as_data(diamond$age)

infected = as_data(diamond$confirmed)
symptomatic= as_data(diamond$symptomatic)
asymptomatic= as_data(diamond$asymptomatic)
deaths = as_data(diamond$deaths)


# Define the prior for the theta parameter
#theta_infected = beta(shape1 = 1, shape2 = 1, dim = categories)
#theta_symptomatic = beta(shape1 = 1, shape2 = 1, dim = categories)
#theta_asymptomatic = beta(shape1 = 1, shape2 = 1, dim = categories)

eta_deaths = lognormal(0, 20)

rho_mean = 20; rho_sd = 5
rho_deaths = lognormal(log(rho_mean/sqrt(1+rho_sd^2/rho_mean^2)), sdlog = sqrt(log(1+rho_sd^2/rho_mean^2)))

# kernel & GP
kernel = rbf(rho_deaths, eta_deaths) 
f = gp(ages, kernel)  
theta_deaths = ilogit(f) # convert to unit interval

# Define the likehood for the model
#distribution(infected) = binomial(size = exposure, prob = theta_infected, dim = categories)  
#distribution(symptomatic) = binomial(size = exposure, prob = theta_symptomatic, dim = categories)  
#distribution(asymptomatic) = binomial(size = exposure, prob = theta_asymptomatic, dim = categories)
distribution(deaths) = binomial(size = infected, prob = theta_deaths, dim = categories)  

```

We can now establish and compile the model:

```{r results="hide"}
# Establish the model
#m_gp= model(theta_deaths, eta_deaths, rho_deaths)
m_gp_var= model(theta_deaths, eta_deaths, rho_deaths) 

n_samples = n_samples_base; chains=chains_base; warmup=warmup_base*4
S = n_samples * chains # Total number of simulations
draws_gp_var=mcmc(m_gp_var, n_samples = n_samples, warmup = warmup,chains = chains)

```


```{r}
# diagnostics
theta_names = colnames(draws_gp_var[[1]])
mcmc_trace(draws_gp_var, pars = theta_names[1:10])
mcmc_trace(draws_gp_var, pars = theta_names[11:12])
mcmc_intervals(draws_gp_var, pars = theta_names[1:10])
mcmc_intervals(draws_gp_var, theta_names[11:12])

summary(draws_gp_var)

```


```{r}
theta_post_gp_var = 
  draws_gp_var %>% 
  reduce(rbind) %>%
  as_tibble() %>%
  select(-eta_deaths, -rho_deaths) 
  
colnames(theta_post_gp_var) = 
  paste0(rep(c( "theta_deaths_"),
             each=categories) , 1:categories)

theta_post_gp_var = as_tibble(theta_post_gp_var) %>%
  mutate(S=1:S) %>%
  gather(posterior, value, -S) %>%
  mutate(risk_cat = as.factor(gsub(pattern = "_([^_]*)$","",(posterior))), 
           age_cat = as.factor((as.numeric(extract_numeric(posterior)))*10-5))

theta_post_gp_var_sum = theta_post_gp_var %>%
  group_by(age_cat) %>%
  summarize(mean = mean(value),
            low = quantile(value, 0.1),
            high = quantile(value, 0.9)
            ) %>%
  gather(range, value, -mean, -age_cat)

ggplot(theta_post_gp_var_sum) +
  geom_point(aes(age_cat, mean, color = age_cat),size=2) +
  geom_line(aes(age_cat, value, group = age_cat)) +
  labs(title="Diamond Princess: Risk of Death\n with 80% Credible Interval", color="", y="Fatality Rate", x="Age") + 
  theme_bw()


# ggplot(theta_post_gp_var) +
#   geom_violin(aes(age_cat, value, fill = age_cat), alpha=0.5, draw_quantiles = c(0.1, 0.5, 0.9), scale = "width") +
#   labs(title="Diamond Princess: Risk of Death", color="", y="Fatality Rate", x="Age") + 
#   ylim(0,.2)+
#   theme_bw()

gp_par = 
  draws_gp_var %>% 
  reduce(rbind) %>%
  as_tibble() %>%
  select(eta_deaths, rho_deaths) %>%
  gather(par, value)

gp_par_plot = ggplot(gp_par) +
  geom_density(aes(value, fill=par), alpha=0.5) +
  labs(title="GP Pars") + 
  facet_wrap(~par, scales = "free")+
  theme_bw()
gp_par_plot

```

```{r}
risk_band = theta_post_gp_var %>%
  group_by(age_cat) %>%
  summarize(mortality_mean = mean(value),
            mortality_low = quantile(value, probs = .1),
            mortality_high = quantile(value, probs = .9))

gp_mortality = ggplot(risk_band %>% mutate(age= as.numeric(as.character(age_cat)))) +
  geom_line(aes(age, mortality_mean)) + 
  geom_ribbon(aes(age, ymin=mortality_low, ymax=mortality_high), fill= "blue", alpha=.2)+
  theme_bw() + 
  labs(title="GP Model - Fatality Rate: 80% Credible Interval")+
  scale_x_continuous(breaks = diamond$age)
gp_mortality

risk_band

```



# Summary of Results

## A. Combined Age Category Model

### Risk Rates

The theta (or $\theta$) amounts represent the parameters of the binomial distribution and are the rates or infection of fatality.

```{r echo=FALSE}
combined_risk_rates
```


## B. Independent Age Category Model

### Risk Rates

```{r echo=FALSE}
vector_risk_rates
vector_mortality_rates
```


Notes: There are very limited data for the age categories 5, 15 and 95.  Here the uniform plays an outsize role, the resultant posterior values for these age buckets suggests a uniform prior might not be the most optimal prior, and a prior skewed towards lower risk rates might be more appropriate.

## C. Gaussian Process Model: Fixed Rho

Fixed $\rho=10$

### Fatality Rates

```{r echo=FALSE}
gp_mortality_fixed_rho
```


## C. Gaussian Process Model: Variable Rho

Fixed $\rho=10$

### Fatality Rates

```{r echo=FALSE}
gp_mortality
```

### Model Parameters


```{r echo=FALSE}
gp_par_plot
```