---
title: 'Coronavirus Time Series Analysis: Are We Italy?'
output:
  html_document:
    df_print: paged
  word_document: default
  html_notebook: default
---

# Introduction

In analyzing the growth trajectories of the Coronavirus, a key question is whether the US and other countries are going to follow the explosion of cases and deaths in Italy or are going to level out, as is being experienced in China and South Korea.

My view is that in an exponential growth scenario (which we are experiencing), it is the *growth rate* that is all important rather than the nominal number of confirmed cases or deaths. This is what this analysis focuses on. A country with low infected numbers but a high growth rate will quickly overtake a country with high infections but low growth rate.  For example, compare Italian vs. Chinese deaths below.

The growth rate will necessary and eventually convert from an exponential curve to a logistic curve (an "S"-shaped curve) as infection rates saturate a population. However, we are currently nowehere near the saturation point, and so, in the near term, a reduction in growth rate will have to be be driven by other sources.

The rest of this report compares the growth rate in confirmed Covid-19 cases and deaths for various countries.  I fit a Gaussian Process curve to the data to smooth out the volatility and determine a mean for the growth rates.  In particular, I have reviewed the "aligned" growth in confirmed cases and deaths. That is, for each country I aligned the data to start: 

* For confirmed case analysis: The day when 200 cases were confirmed in each respective country
* For deaths analysis: The day when 10 deaths were confirmed in each respective country


The current date of this analysis is March 18, 2020.

All code and data can be found on [Github](https://github.com/howardnewyork/corona).

# Summary of Results

The data has some good news.  China and South Korea have managed to remarkably bring down the growth of new infections and deaths.  Other countries do not have such good news and the confirmed cases and deaths are doubling every four or five days.  Although Italy is experiencing a total disaster, it appears as if the severe exponential growth it was experiencing is slowing a little. The growth rates in Italy are still very high but trending in the right direction.  Let's see if this trend will continue.

For US and other European countries, the news is uniformally bad. These countries seem to be following more the Italian model rather than the South Korean one.


I note that the Chinese data is so good that it seems suspect. They had only 25 new infections and 13 deaths on March 19.  Given their early problems and wide exposure, the new data seem too good to be true.

In summary, there is an inkling of good news in the Italian data and hopefully the trend continues.  South Korea and China (if the data are correct) show that the virus can be beat, BUT at this stage radical control of population movement is likely necessary (per Italian and Chinese model).  


# Methodology

I used a Gaussian Process model to establish the mean growth rates.  Raw growth rates were determined as:

* Confirmed case growth rate:  $Confirmed~Case~ Growth~Rate_t = log(Confirmed~Cases_{t} / Confirmed~Cases_{t-1} )$
* Confirmed death growth rate:  $Deaths~Case~ Growth~Rate_t = log(Confirmed~Deaths_{t} / Confirmed~Deaths_{t-1} )$

I then calculate a 95% credible around such mean estimate.



# Data Sources

[Johns Hopkins COID-19 Data Repository](https://github.com/CSSEGISandData/COVID-19/compare?expand=1)

Note: the data are very messy and  dependent on differing testing regimes rather than true infection experience.


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8, warnings= FALSE)
```

```{r echo=FALSE, message=FALSE}
# Time Series Analysis of Survival from Breast Cancer


library(rstan)
options(mc.cores = min(8,parallel::detectCores()))
rstan_options(auto_write = TRUE)

library(ggplot2)
library(dplyr)
library(purrr)
library(tidyr)
library(readr)
library(stringr)
library (bayesplot)
library(lubridate)
library(corrplot)

# MCMC Pars
iter = 350
warmup = 150
chains = 8



# Set Directories
working_dir = paste0(getwd(), "/")
data_dir = paste0(working_dir, "data/")
output_dir = paste0(working_dir, "output/")
stan_dir = paste0(working_dir, "stan/")

#parameters
selected_countries = c("Germany", "France", "Spain", "China", "Australia", "US", "South Korea", "Italy",  "United Kingdom")
cutoff= 200
cutoff_deaths= 10
file_date = "2020_03_24"

# Load Data for Confirmed Cases
#corona_raw = read_csv(file = paste0(data_dir, paste0("time_series_19_covid_", file_date, ".csv")))
corona_raw = read_csv(file = paste0(data_dir, "time_series_covid19_confirmed_global.csv")) %>%
  rename(province_state = "Province/State",
         country = "Country/Region",
         lat = Lat,
         long = Long)

corona_raw[1:5,1:5]

corona = corona_raw %>%
  mutate(country = if_else(country == "Korea, South", "South Korea", country))  %>%  
  select(-lat, -long, -province_state) %>%
  group_by(country) %>%
  summarise_all(sum) %>%
  gather(date, confirmed, -country) %>%
  group_by(country) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(growth = confirmed - lag(confirmed),
         growth_rate = log(confirmed / lag(confirmed))) %>%
  arrange(date) %>%
  mutate(growth_rate = ifelse(is.finite(growth_rate), growth_rate, NA))  


corona_align = corona %>%
  filter(country %in% selected_countries,
         confirmed >=cutoff) %>%
  ungroup() %>%
  group_by(country) %>%
  mutate(day = date-first(date)+1) %>%
  select(country,confirmed, growth,growth_rate, date, day) %>%
  arrange(day) %>%
  remove_missing()

#deaths_raw = read_csv(file = paste0(data_dir, paste0("time_series_19_covid_deaths_", file_date, ".csv")))
deaths_raw = read_csv(file = paste0(data_dir,  "time_series_covid19_deaths_global.csv")) %>%
  rename(province_state = "Province/State",
         country = "Country/Region",
         lat = Lat,
         long = Long)

deaths = deaths_raw %>%
  mutate(country = if_else(country == "Korea, South", "South Korea", country))  %>%  
  select(-lat, -long, -province_state) %>%
  group_by(country) %>%
  summarise_all(sum) %>%
  gather(date, confirmed, -country) %>%
  group_by(country) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(growth = confirmed - lag(confirmed),
         growth_rate = log(confirmed / lag(confirmed))) %>%
  arrange(date) %>%
  mutate(growth_rate = ifelse(is.finite(growth_rate), growth_rate, NA))  


deaths_align = deaths %>%
  filter(country %in% selected_countries,
         confirmed >=cutoff_deaths) %>%
  ungroup() %>%
  group_by(country) %>%
  mutate(day = date-first(date)+1) %>%
  select(country,confirmed, growth,growth_rate, date, day) %>%
  arrange(day) %>%
  remove_missing()


```



# Data Exploration


```{r echo=FALSE}
ggplot(corona %>% filter(country %in% selected_countries)) +
  geom_line(aes(date,confirmed, color=country)) + 
  geom_point(aes(date,confirmed, color=country, shape=country)) + 
  labs(title = paste(c("Confirmed Cases by Country\n",file_date), collapse=""), x="Date", y="Confirmed Cases")+
   scale_y_continuous( labels = scales::comma) +
  theme_bw()

if (F){
ggplot(corona %>% filter(country %in% selected_countries)) +
  geom_line(aes(date,growth, color=country)) + 
  labs(title = paste(c("Growth in Confirmed Cases by Country\n",file_date), collapse=""), x="Date", y="Growth in Confirmed Cases")+
   scale_y_continuous( labels = scales::comma)

corona %>%  filter(country %in% selected_countries, confirmed >=10) %>% select(date, confirmed, growth_rate, growth_rate) %>% 

ggplot() +
  geom_line(aes(date,growth_rate, color=country)) + 
  labs(title = paste(c("Growth in Confirmed Cases by Country", "\n",file_date), collapse=""), x="Date", y="Daily Growth Rate")+
  scale_y_continuous( labels = scales::comma) +
  theme_bw() +
  coord_cartesian(ylim=c(-1, 3))
}


```

```{r echo=FALSE}


ggplot(corona_align ) +
  geom_line(aes(day,confirmed, color=country)) + 
  geom_point(aes(day,confirmed, color=country, shape=country)) + 
  labs(title = paste(c("Aligned Confirmed Cases by Country\n",file_date), collapse=""), x=paste0(c("days after ", cutoff, " Cases"), collapse = ""), y="Confirmed Cases")+
   scale_y_continuous( labels = scales::comma) + theme_bw()

if (F){
ggplot(corona_align ) +
  geom_line(aes(day,growth, color=country)) + 
  geom_point(aes(day,growth, color=country, shape=country)) + 
  labs(title = paste(c("Aligned Growth in Confirmed Cases by Country\n",file_date), collapse=""), x=paste0(c("days after ", cutoff, " Cases"), collapse=""), y="Growth in Confirmed Cases")+
   scale_y_continuous( labels = scales::comma)

ggplot(corona_align) +
  geom_line(aes(day,growth_rate, color=country)) + 
  geom_point(aes(day,growth_rate, color=country, shape=country)) + 
  labs(title = paste(c("Aligned Growth Rate by Country\n",file_date), collapse=""), x=paste0(c("days after ", cutoff, " Cases"), collapse=""), y="Daily Growth Rate")+
  theme_bw() 
corona_cor = corona_align %>%
  select(day, country, growth_rate) %>%
  pivot_wider(names_from = country, values_from = growth_rate) %>%
  select(-day) %>% 
  cor(use = "pairwise.complete.obs")
corrplot(corona_cor[selected_countries, selected_countries])
corona_align %>%
  group_by(country) %>%
  group_map(~lm(data = .x,  formula =  growth_rate ~ day ))


ave_growth = corona_align %>%
  group_by(country) %>%
  summarize(mean_growth = mean(growth_rate, na.rm = T),
            median_growth = median(growth_rate, na.rm = T) )

ggplot(ave_growth %>% pivot_longer(names_to = "type", values_to = "value", -country)) + 
  geom_point(aes(country, value, color=type), size=2)+ 
  labs(title = paste(c("Mean and Median Growth\n",file_date), collapse="")) + 
  theme_bw()

}




if (F){

corona_align = corona_align %>%
  filter(is.finite(growth_rate)) %>%
  mutate(smoothed = predict(loess(formula =  growth_rate ~ as.numeric(day), na.action = na.omit)))

corona_align


ggplot(corona_align) +
  geom_line(aes(day,smoothed, color=country)) + 
  geom_point(aes(day,smoothed, color=country, shape=country)) + 
  labs(title = paste("Smoothed Aligned Growth Rate by Country\n",file_date, collapse=""), x=paste0(c("days after ", cutoff, " Cases"), collapse=""), 
       y="Daily Growth Rate")+
  theme_bw() 

corona_cor2 = corona_align %>%
  select(day, country, smoothed) %>%
  pivot_wider(names_from = country, values_from = smoothed) %>%
  select(-day) %>% 
  cor(use = "pairwise.complete.obs")
corrplot(corona_cor2[selected_countries, selected_countries])

#corr2 = corona_cor2
  
}

```


```{r echo=FALSE}
ggplot(deaths %>% filter(country %in% selected_countries)) +
  geom_line(aes(date,confirmed, color=country)) + 
  geom_point(aes(date,confirmed, color=country, shape=country)) + 
  labs(title = paste(c("Deaths by Country\n",file_date), collapse=""), x="Date", y="Confirmed Cases")+
   scale_y_continuous( labels = scales::comma) + 
  theme_bw()


ggplot(deaths_align ) +
  geom_line(aes(day,confirmed, color=country)) + 
  geom_point(aes(day,confirmed, color=country, shape=country)) + 
  labs(title = paste(c("Aligned Deaths by Country\n",file_date), collapse=""), x=paste0(c("days after ", cutoff_deaths, " Deaths"), collapse = ""), y="Total Deaths")+
   scale_y_continuous( labels = scales::comma) + 
  theme_bw()

if(F){
ave_growth_deaths = deaths_align %>%
  group_by(country) %>%
  summarize(mean_growth = mean(growth_rate, na.rm = T),
            median_growth = median(growth_rate, na.rm = T) )

ggplot(ave_growth_deaths %>% pivot_longer(names_to = "type", values_to = "value", -country)) + 
  geom_point(aes(country, value, color=type), size=2)+ 
  labs(title = paste(c("Mean and Median Daily Growth in Deaths\n",file_date), collapse=""), y= "Daily Growth Rate", x="Country") + 
  theme_bw()
}

```


# Growth Rate Analysis

```{r echo=FALSE}
country_count = length(unique(corona_align$country))
N = nrow(corona_align)
N_pred = as.numeric(country_count * (max(corona_align$day) - min(corona_align$day)+1))


stan_list = list(
  N1 = N,  # historic data
  #N2= N_pred,  # predicted points
  N2= N,  # predicted points
  D=1, # Dimension of input values
  dimCat = country_count, # dimension of number o f unique categorical values
  
  # Historic Data
  x1 = matrix(as.numeric(corona_align$day), ncol = 1) ,  # input values (N1 x D matrix from R)
  x_cat1= as.numeric(as.factor(corona_align$country)),  # input values containing a categorical value
  y1 = corona_align$growth_rate,  # output values 
  # Inputs for prediction
  
  # x2= matrix(rep(min(corona_align$day):max(corona_align$day), each = country_count), ncol=1),  # input values (N xD matrix from R)
  # x_cat2 =  rep(1:country_count, times = max(corona_align$day)),  # input values containing a categorical value
  x2 = matrix(as.numeric(corona_align$day), ncol = 1) ,  # input values (N1 x D matrix from R)
  x_cat2= as.numeric(as.factor(corona_align$country)),  # input values containing a categorical value

 # inputs for priors on parameters
 prior_eta = 2,
 prior_rho = c(5,5),
 prior_sigma = 1,
 prior_alpha =2
)
```



```{r echo=FALSE}
country_count = length(unique(deaths_align$country))
N = nrow(deaths_align)
N_pred = as.numeric(country_count * (max(deaths_align$day) - min(deaths_align$day)+1))


stan_list_deaths = list(
  N1 = N,  # historic data
  #N2= N_pred,  # predicted points
  N2= N,  # predicted points
  D=1, # Dimension of input values
  dimCat = country_count, # dimension of number oif unique categorical values
  
  # Historic Data
  x1 = matrix(as.numeric(deaths_align$day), ncol = 1) ,  # input values (N1 x D matrix from R)
  x_cat1= as.numeric(as.factor(deaths_align$country)),  # input values containing a categorical value
  y1 = deaths_align$growth_rate,  # output values 
  # Inputs for prediction
  
  # x2= matrix(rep(min(deaths_align$day):max(deaths_align$day), each = country_count), ncol=1),  # input values (N xD matrix from R)
  # x_cat2 =  rep(1:country_count, times = max(deaths_align$day)),  # input values containing a categorical value
  x2 = matrix(as.numeric(deaths_align$day), ncol = 1) ,  # input values (N1 x D matrix from R)
  x_cat2= as.numeric(as.factor(deaths_align$country)),  # input values containing a categorical value

 # inputs for priors on parameters
 prior_eta = 2,
 prior_rho = c(5,5),
 prior_sigma = 1,
 prior_alpha =2
)
```


##  Confirmed Cases

```{r echo=FALSE}
mod_block = stan_model(file = paste0(stan_dir, "gp_mod2.stan"))

draws_block = sampling(object = mod_block, data = stan_list , chains = chains, iter = iter, warmup=warmup)
#summary(draws_block, pars = c("eta", "rho", "sigma"))$summary
```


```{r echo=FALSE}

country_names = levels(as.factor(corona_align$country))

cap_day = 360
cap_rate = log(2) / 360
preds_block = tibble(
  day = c(stan_list$x2),
  country = country_names[stan_list$x_cat2],
  growth_rate = summary(draws_block, pars = c("fStar"))$summary[,"mean"],
  gr_low = summary(draws_block, pars = c("fStar"))$summary[,"2.5%"],
  gr_high = summary(draws_block, pars = c("fStar"))$summary[,"97.5%"],
  double_days = ifelse(growth_rate<=cap_rate, NA, log(2) /growth_rate),
  double_low = ifelse(growth_rate<=cap_rate, NA, log(2) /gr_low),
  double_high = ifelse(growth_rate<=cap_rate, NA,log(2) /gr_high)) 

preds_block = preds_block %>%
  mutate(
    gr_low = pmax(0, gr_low),
    gr_high = pmax(0, gr_high),
    double_days = pmax(0, double_days),
    double_low = pmax(0,double_low),
    double_high = pmax(0,double_high)
    #double_low = ifelse(double_days == cap_day, cap_day, double_low),
    #double_high = ifelse(double_days == cap_day, cap_day, double_high)
  )



(plot_rate_corona = ggplot(preds_block)  +
  geom_line(aes(day, growth_rate, color=country)) + 
  geom_ribbon(aes(day, ymin=gr_low,ymax=gr_high, fill=country), alpha=0.2) + 
  geom_point(aes(day, growth_rate, color=country, shape=country))  +
  labs(title = paste(c("Daily Growth Rate of Infection following\n", cutoff, " confirmed infections\n", file_date), collapse =""), x=paste0(c("days after ", cutoff, " Cases"), collapse=""), y="Mean Field Growth Rate") + 
  #coord_cartesian(ylim=c(0, .5))+
  facet_wrap(~country, scales="free") + 
  theme_bw())
  


(plot_double_corona = ggplot(preds_block)  +
  geom_line(aes(day, double_days, color=country)) + 
  geom_ribbon(aes(day, ymin=double_low,ymax=double_high, fill=country), alpha=0.2) + 
  geom_point(aes(day, double_days, color=country, shape=country))  +
  labs(title = paste0(c("Days until Confirmed Infection Count is Doubled\n Capped at ", cap_day," Days\n",file_date),collapse = ""), 
       x=paste0(c("days after ", cutoff, " Confirmed"), collapse=""), y="Days to Double Confirmed") + 
  facet_wrap(~country, scales="free") + 
  theme_bw())
  
ggsave(file = paste0(c(output_dir, "plot_rate_corona_",file_date,".png"), collapse = ""), plot = plot_rate_corona)
ggsave(file = paste0(c(output_dir, "plot_double_corona_",file_date,".png"), collapse = ""), plot = plot_double_corona)

```






## Deaths

```{r echo=FALSE}

draws_block_deaths = sampling(object = mod_block, data = stan_list_deaths , chains = chains, iter = iter, warmup=warmup)
```


```{r echo=FALSE }

country_names = levels(as.factor(deaths_align$country))

cap_day = 90
preds_block_deaths = tibble(
  day = c(stan_list_deaths$x2),
  country = country_names[stan_list_deaths$x_cat2],
  growth_rate = summary(draws_block_deaths, pars = c("fStar"))$summary[,"mean"],
  gr_low = summary(draws_block_deaths, pars = c("fStar"))$summary[,"2.5%"],
  gr_high = summary(draws_block_deaths, pars = c("fStar"))$summary[,"97.5%"],
  double_days = ifelse(log(2) /growth_rate>cap_day, cap_day, log(2) /growth_rate),
  double_low = ifelse(log(2) /gr_low>cap_day,cap_day, log(2) /gr_low),
  double_high = ifelse(log(2) /gr_high>cap_day,cap_day, log(2) /gr_high)) 

preds_block_deaths = preds_block_deaths %>%
  mutate(
    gr_low = pmax(0, gr_low),
    gr_high = pmax(0, gr_high),
    double_days = pmax(0, double_days),
    double_low = pmax(0,double_low),
    double_high = pmax(0,double_high),
    double_low = ifelse(double_days == cap_day, cap_day, double_low),
    double_high = ifelse(double_days == cap_day, cap_day, double_high)
  ) 



(plot_rate_death = ggplot(preds_block_deaths)  +
  geom_line(aes(day, growth_rate, color=country)) + 
  geom_ribbon(aes(day, ymin=gr_low,ymax=gr_high, fill=country), alpha=0.2) + 
  geom_point(aes(day, growth_rate, color=country, shape=country))  +
  labs(title = paste(c("Daily Smoothed Log Growth Rate of Deaths following\n", cutoff_deaths, " Confirmed Deaths (Rate = log(Cum. Deaths(t)/Cum. Deaths(t-1))\n", file_date), collapse =""), x=paste0(c("days after ", cutoff_deaths, " Deaths"), collapse=""), y="Mean Field Growth Rate") + 
  #coord_cartesian(ylim=c(0, .5))+
  facet_wrap(~country, scales="free") + 
  theme_bw())
  
(plot_double_death = ggplot(preds_block_deaths)  +
  geom_line(aes(day, double_days, color=country)) + 
  geom_ribbon(aes(day, ymin=double_low,ymax=double_high, fill=country), alpha=0.2) + 
  geom_point(aes(day, double_days, color=country, shape=country))  +
  labs(title = paste0(c("Days until Death Count is Doubled\n Capped at ", cap_day," Days\n",file_date), collapse = ""), 
       x=paste0(c("days after ", cutoff_deaths, " Deaths"), collapse=""), y="Days to Double Deaths") + 
  facet_wrap(~country, scales="free") + 
  theme_bw())
  

ggsave(file = paste0(c(output_dir, "plot_rate_death_",file_date,".png"), collapse = ""), plot = plot_rate_death)
ggsave(file = paste0(c(output_dir, "plot_double_death_",file_date,".png"), collapse = ""), plot = plot_double_corona)


```




